Division and Factorization of p-Adic Polynomials

### 2016-01-08 polynomials/muldivfact.tex

\begin{deftn}
\label{def:nondeg}
Let $P = P_0 + O(\varphi_P)$. We say that the Newton precision 
$O(\varphi_P)$ on $P$ is \emph{nondegenerate} if $\varphi_P \geq 
\NF(P_0)$ and $\varphi_P(x) > y$ for all extremal point $(x,y)$ of 
$\NP(P_0)$.
\end{deftn}

We notice that, under the conditions of the above definition, the
Newton polygon of $P$ is well defined. Indeed, if $\delta P$ is any
polynomial whose Newton function is not less than $\varphi_P$, we
have $\NP(P_0 + \delta P) = \NP(P_0)$.

@@

\noindent
We fix an integer $n$. A lattice precision datum for a polynomial of 
degree $n$ is a lattice $H$ lying in the vector space $K_{\leq n}[X]$. 
We shall sometimes denote it $O(H)$ in order to emphasize that it should 
be considered as a precision datum. The notation $P(X) + O(H)$ then 
refers to any polynomial in the $W$-affine space $P(X) + H$. Tracking
lattice precision can be done using differentials as shown in
\cite[Lemma~3.4 and Proposition~3.12]{padicprec}: if $f : K_{\leq n}[X] 
\to K_{\leq m}[X]$ denotes any strictly differentiable function with
surjective differential, under mild assumption on $H$, we have:
$$f(P(X)+H) = f(P(X)) + f'(P(X))(H)$$
where $f'(P(X))$ denotes the differential of $f$ at $P(X)$. The
equality sign reflets the optimality of the method.

@@

\begin{deftn}
Let $\varphi$ and $\psi$ be two Newton functions of degree $n$ and
$d$ respectively. Set $\mu = \psi(d) - \psi(d-1)$. Letting $\Delta$
be the greatest affine function of slope $\mu$ with $\Delta \leq
\varphi_{|[d,n]}$ and $\delta = \Delta(d) - \psi(d)$, we define:
$$\begin{array}{r@{\,\,}rcl}
\varphi \nfmod \psi & 
\multicolumn{3}{@{}l}{= \varphi_{|[0,d{-}1]} \nfplus \big( 
\psi_{|[0,d{-}1]} + \delta\big)} \medskip \\
\varphi \nfdiv \psi & : \, [0, n-d] & \to & \R \cup \{+\infty\} \smallskip \\
& x & \mapsto & \inf_{h \geq 0} \psi(x+d+h) - \mu h.
\end{array}$$
\end{deftn}

@@

\begin{theo}
\label{theo:EDivisionNP}
Let $A, B \in K[X]$ with $B \neq 0$. Let $d$ be the degree of $B$.
Let $\mu = \NF(B)(d) - \NF(B)(d{-}1)$ be the last slope of $B$ and 
$y = \Delta(x)$ be the
equation of the highest line of slope $\mu$ lying below $\NP(A)$.
Then:
\begin{align}
\NF(A \mod B) & \geq \NF(B) \nfmod \NF(A) \label{eq:AmodB} \\
\text{and} \hspace{4mm}
\NF(A \div B) & \geq \NF(B) \nfdiv \NF(A) \label{eq:AdivB}
\end{align}
\end{theo}

@@

Let us now prove Eq.~\eqref{eq:AmodB}. 
Replacing $B$ by $c^{-1} B$ where $c$ denotes the leading coefficient
of $B$, we may assume that $B$ is monic. Using linearity, we may further
assume that $A$ is a monomial. Set $R_n = X^n \mod B$. The relation
we have to prove is:
$$\NF(R_n)(x) \geq \NF(B)(x) - \mu(n-d)
\quad \text{for } n\geq d.$$
We proceed by induction. The initialisation is clear because $R_d$
agrees with $(-B)$ up to degree $d{-}1$. We have the relation
$R_{n+1} = X R_n - c_n B$
where $c_n$ is the coefficient in $X^{d-1}$ of $R_n$. Thanks to the
induction hypothesis, we have:
\begin{align*}
\val(c_n) \geq \NF(R_n)(d{-}1) & \geq \NF(B)(d{-}1) - \mu(n{-}d) \\
& = - \mu(n{+}1{-}d)
\end{align*}
since $\mu = -\NF(B)(d{-}1)$ because $B$ is monic. Therefore
$\NF(c_n B)(x) \geq \NF(B)(x) - \mu(n{+}1{-}d)$ for all $x$. On the 
other hand, for all $x$, we have:
$$\NF(X R_n)(x) = \NF(R_n)(x{-}1) \geq \NF(R_n)(x) - \mu$$
from what we get $\NF(X R_n)(x) \geq \NF(B)(x) - \mu(n{+}1{-}d)$. As a
consequence
$\NF(R_{n+1})(x) \geq \NF(B)(x) - \mu(n{+}1{-}d)$
and the induction follows.
Eq.~\eqref{eq:AdivB} is now derived from:
$$\NF(A \div B) \nftimes \NF(B) \geq \NF(A) \nfplus \NF(A \mod B)$$
using the estimation on $\NF(A\mod B)$ we have just proved
(see Figure~\ref{fig:NewtonEuclide}).
\end{proof}

@@

\noindent
{\bf Newton precision.}
We consider here the case where the precision is given by Newton
functions. Concretely, we pick $A, B \in K[X]$ two polynomials which 
are known at precision $O(\varphi_A)$ and $O(\varphi_B)$ respectively:
$$A = A_0 + O(\varphi_A)
\quad \text{and} \quad
B = B_0 + O(\varphi_B)$$
Here $A_0$ and $B_0$ are some approximations of $A$ and $B$ respectively 
and $\varphi_A$ and $\varphi_B$ denotes two Newton functions of degree 
$\deg A$ and $\deg B$ respectively.
We are interesting in determining the precision on $A \mod B$ and $A
\div B$. The following proposition gives a theoretical answer under
mild assumptions.

@@

\begin{proof}
Let $\delta A$ (resp. $\delta B$) be a polynomial whose Newton 
function is not less than $\varphi_A$ (resp. $\varphi_B$) and define
$\delta Q$ and $\delta R$ by:
\begin{align*}
Q_0 + \delta Q & = (A_0 + \delta A) \div (B_0 + \delta B) \\
R_0 + \delta R & = (A_0 + \delta A) \mod (B_0 + \delta B)
\end{align*}
where $Q_0 = A_0 \div B_0$ and $R_0 = A_0 \mod B_0$.
We have to show that $\NF(\delta Q) \geq \varphi \nfdiv \NF(B)$ and
$\NF(\delta R) \geq \varphi \nfmod \NF(B)$.
Set $\delta X = \delta A - Q_0 \delta B$. 
Using Theorem~\ref{theo:EDivisionNP}, we obtain $\NF(Q_0) \geq
\NF(A) \nfdiv \NF(B)$ and consequently $\NF(\delta X) \geq \varphi$.
On the other hand, an easy computation yields
$$\delta X = (B + \delta B) \cdot \delta Q + \delta R$$
so that $\delta Q = \delta X \div (B + \delta B)$ and 
$\delta R = \delta X \mod (B + \delta B)$. Using again 
Theorem~\ref{theo:EDivisionNP}, we get the desired result.
\end{proof}

@@

\noindent
{\bf Lattice precision.}
We now move to lattice precision. We pick $A$ and $B$ two polynomials 
of respective degree $n$ and $d$ and assume that they are known at 
precision $O(H_A)$ and $O(H_B)$ respectively:
$$A = A_0 + O(H_A)
\quad \text{and} \quad
B = B_0 + O(H_B).$$
where $H_A \in K_{\leq n}[X]$ and $H_B \in K_{\leq d}[X]$ are lattices. 
According to the results of \cite{padicprec}, in order to determine the 
precision on $A \div B$ and $A \mod B$, we need to compute the 
differential of the mappings $(X,Y) \mapsto X \div Y$ and $(X,Y)
\mapsto X \mod Y$ at the point $(A_0, B_0)$. Writing $Q_0 = A_0 \div
B_0$ and $R_0 = A_0 \mod B_0$, this can be done by
expanding the relation:
$$A_0 + dA = (B_0 + dB) (Q_0 + dQ) + (R_0 + dR)$$
and neglecting the terms of order $\geq 2$. We get this way 
$dA = B_0 dQ + Q_0 dB + dR$
meaning that $dQ$ and $dR$ appears respectively as the quotient and
the remainder of the Euclidean division of $dX = dA - B_0 dQ$ by $B_0$.

@@

For this example, we work over $W = \Z_2$ and fix a monic polynomial $M 
\in \Z[X]$ (known exactly) of degree $5$. Our aim is to compare the 
numerical stability of the multiplication in the quotient $\Z_2[X]/M(X)$ 
depending on the precision model we are using. In order to do so, we 
pick $n$ random polynomials $P_1, \ldots, P_n$ in $\Z_2[X]/M(X)$ 
(according to the Haar measure) whose coefficients are all known at 
precision $O(2^N)$ for some large integer $N$. We then compute the 
product of the $P_i$'s using the following quite naive algorithm.

@@

\noindent $\bullet$
in the case of jagged and Newton precision, the precision on the output 
may be written into the form $\sum_{i=1}^4 O(2^{N_i}) \: X^i$ and 
$G = \sum_{i=1}^4 (N_i - N)$;

@@

We define the sequences $(A_i)$ and $(V_i)$ by the recurrence: 
\begin{align*} 
A_0 & = A_\app, \quad V_{0} = V_\app \medskip \\ A_{i+1} 
&= A_i + \big(V_i (P_\app + \delta_P) \:\mod\: A_i\big), \smallskip \\ 
V_{i+1} &= (2 V_i -V_i^2 B_{i+1} ) \mod A_{i+1} \\ 
&  \hspace{1cm}\text{where } B_{i+1} = (P_\app + \delta_P) \div A_{i+1}. 
\end{align*} 
where $A_\app$ and $V_\app$ are those related to $P_\app$. Note that 
$A_\app = a_d \cdot A^{(1)}_\app$ if $a_d$ denotes the coefficient of 
$X^d$ in $P$. By Remark~\ref{rem:slope-factor}, we know that the 
sequence $(A_i)$ converges to $A = a_d \cdot A^{(1)}$ and furthermore:
$$\NP(A - A_\app) \geq \varphi_{[0,d{-}1]} + 
b_\varphi\big((V_\app \cdot \delta P) \mod A_\app\big)$$ 
since $(V_\app P_\app) \mod A_\app = (1 - V_\app B_\app) \mod A_\app = 
0$. Using repeatedly Lemma~\ref{lem:bphib0}, we obtain:
$$b_\varphi\big((V_\app \cdot \delta P) \mod A_\app\big)
  \geq b_0(V_\app) + b_\varphi(\delta P) \\
  \geq b_\varphi(\delta P) \geq \delta.$$
Thus $\NP(A - A_\app) \geq \varphi_{[0,d{-}1]} + \delta$. Dividing
by $a_d$, we find
$\NP(A^{(1)} - A^{(1)}_\app) \geq \varphi_{A^{(1)}}$ and we are done.
\end{proof}

@@

\noindent
{\bf Lattice precision.}
In the case of lattice precision, the answer to the first question is 
then given by the theory of \cite{padicprec} and is controlled by the 
differential of the application sending $P$ to $A^{(1)}$.

\begin{prop}
\label{prop:precA1}
The application $P \mapsto A^{(1)}$ (where $A^{(1)}$ is defined as before) is 
of class $C^1$ in a neighborhood of $P_0$ and its differential 
at $P_0$ is the linear mapping:
$$dP \mapsto dA^{(1)} = (V^{(1)} \: dP) \mod A^{(1)}$$
where $V^{(1)}$ is the inverse of $B^{(1)}$ modulo $A^{(1)}$.
\end{prop}

\begin{proof}
There certain exists a neighborhood $\mathcal U$ of $P_0$ on which the 
Newton polygon does not vary. On $\mathcal U$, the function $f : P 
\mapsto (A^{(1)},B^{(1)})$ is well defined. It is furthermore bijective 
and its inverse $g$ is given by $(A^{(1)},B^{(1)}) \mapsto A^{(1)}B^{(1)}$.
Clearly $g$ is of class $C^1$ and its differential is given by
\begin{equation}
\label{eq:dPdAdB}
(dA^{(1)}, dB^{(1)}) \mapsto dP = A^{(1)} \: dB^{(1)} + B^{(1)} \: dA^{(1)}.
\end{equation}
Thanks to Bézout Theorem, it is invertible as soon as $A^{(1)}$ and $B^{(1)}$ are
coprime. This happens on $f(\mathcal U) = g^{-1}(\mathcal U)$ because 
$\NP(A^{(1)})$ and $\NP(B^{(1)})$ do not shape a common slope. As a consequence
$f$ is of class $C^1$ on $\mathcal U$. Moreover, its differential is
obtained by inverting Eq.~\eqref{eq:dPdAdB}. Reducing modulo $A^{(1)}$, we
get $dP \equiv B^{(1)} \: dA^{(1)} \pmod {A^{(1)}}$.
The claimed result follows after having noticed that $dA^{(1)}$
has degree at most $d{-}1$.
\end{proof}

>>>>

\begin{deftn}
\label{def:nondeg}
Let $P = P_\app + O(\varphi_P)$. We say that the Newton precision 
$O(\varphi_P)$ on $P$ is \emph{nondegenerate} if $\varphi_P \geq 
\NF(P_\app)$ and $\varphi_P(x) > y$ for all extremal point $(x,y)$ of 
$\NP(P_\app)$.
\end{deftn}

We notice that, under the conditions of the above definition, the
Newton polygon of $P$ is well defined. Indeed, if $\delta P$ is any
polynomial whose Newton function is not less than $\varphi_P$, we
have $\NP(P_\app + \delta P) = \NP(P_\app)$.

@@

\noindent
We fix an integer $n$. A lattice precision datum for a polynomial of 
degree $n$ is a lattice $H$ lying in the vector space $K_{\leq n}[X]$. 
We shall sometimes denote it $O(H)$ in order to emphasize that it should 
be considered as a precision datum. The notation $P_\app(X) + O(H)$ then 
refers to any polynomial in the $W$-affine space $P_\app(X) + H$. 
Tracking lattice precision can be done using differentials as shown in
\cite[Lemma~3.4 and Proposition~3.12]{padicprec}: if $f : K_{\leq n}[X] 
\to K_{\leq m}[X]$ denotes any strictly differentiable function with
surjective differential, under mild assumption on $H$, we have:
$$f(P_\app(X)+H) = f(P_\app(X)) + f'(P_\app(X))(H)$$
where $f'(P_\app(X))$ denotes the differential of $f$ at $P_\app(X)$. 
The equality sign reflets the optimality of the method.

@@

\begin{deftn}
Let $\varphi$ and $\psi$ be two Newton functions of degree $n$ and
$d$ respectively. Set $\lambda = \psi(d) - \psi(d-1)$. Letting $\Delta$
be the greatest affine function of slope $\lambda$ with $\Delta \leq
\varphi_{|[d,n]}$ and $\delta = \Delta(d) - \psi(d)$, we define:
$$\begin{array}{r@{\,\,}rcl}
\varphi \nfmod \psi & 
\multicolumn{3}{@{}l}{= \varphi_{|[0,d{-}1]} \nfplus \big( 
\psi_{|[0,d{-}1]} + \delta\big)} \medskip \\
\varphi \nfdiv \psi & : \, [0, n-d] & \to & \R \cup \{+\infty\} \smallskip \\
& x & \mapsto & \inf_{h \geq 0} \psi(x+d+h) - \lambda h.
\end{array}$$
\end{deftn}

@@

\begin{theo}
\label{theo:EDivisionNP}
Given $A, B \in K[X]$ with $B \neq 0$, we have:
\begin{align}
\NF(A \mod B) & \geq \NF(A) \nfmod \NF(B) \label{eq:AmodB} \\
\text{and} \hspace{4mm}
\NF(A \div B) & \geq \NF(A) \nfdiv \NF(B) \label{eq:AdivB}
\end{align}
\end{theo}

@@

Let us now prove Eq.~\eqref{eq:AmodB}. 
Replacing $B$ by $c^{-1} B$ where $c$ denotes the leading coefficient
of $B$, we may assume that $B$ is monic. Using linearity, we may further
assume that $A$ is a monomial. Set $R_n = X^n \mod B$. The relation
we have to prove is:
$$\NF(R_n)(x) \geq \NF(B)(x) - \lambda(n-d)
\quad \text{for } n\geq d.$$
We proceed by induction. The initialisation is clear because $R_d$
agrees with $(-B)$ up to degree $d{-}1$. We have the relation
$R_{n+1} = X R_n - c_n B$
where $c_n$ is the coefficient in $X^{d-1}$ of $R_n$. Thanks to the
induction hypothesis, we have:
\begin{align*}
\val(c_n) \geq \NF(R_n)(d{-}1) & \geq \NF(B)(d{-}1) - \lambda(n{-}d) \\
& = - \lambda(n{+}1{-}d)
\end{align*}
since $\lambda = -\NF(B)(d{-}1)$ because $B$ is monic. Therefore
$\NF(c_n B)(x) \geq \NF(B)(x) - \lambda(n{+}1{-}d)$ for all $x$. On the 
other hand, for all $x$, we have:
$$\NF(X R_n)(x) = \NF(R_n)(x{-}1) \geq \NF(R_n)(x) - \lambda$$
from what we get $\NF(X R_n)(x) \geq \NF(B)(x) - \lambda(n{+}1{-}d)$. As a
consequence
$\NF(R_{n+1})(x) \geq \NF(B)(x) - \lambda(n{+}1{-}d)$
and the induction follows.
Eq.~\eqref{eq:AdivB} is now derived from:
$$\NF(A \div B) \nftimes \NF(B) \geq \NF(A) \nfplus \NF(A \mod B)$$
using the estimation on $\NF(A\mod B)$ we have just proved
(see Figure~\ref{fig:NewtonEuclide}).
\end{proof}

@@

\noindent
{\bf Newton precision.}
We consider here the case where the precision is given by Newton
functions. Concretely, we pick $A, B \in K[X]$ two polynomials which 
are known at precision $O(\varphi_A)$ and $O(\varphi_B)$ respectively:
$$A = A_\app + O(\varphi_A)
\quad \text{and} \quad
B = B_\app + O(\varphi_B)$$
Here $A_\app$ and $B_\app$ are some approximations of $A$ and $B$ respectively 
and $\varphi_A$ and $\varphi_B$ denotes two Newton functions of degree 
$\deg A$ and $\deg B$ respectively.
We are interesting in determining the precision on $A \mod B$ and $A
\div B$. The following proposition gives a theoretical answer under
mild assumptions.

@@

\begin{proof}
Let $\delta A$ (resp. $\delta B$) be a polynomial whose Newton 
function is not less than $\varphi_A$ (resp. $\varphi_B$) and define
$\delta Q$ and $\delta R$ by:
\begin{align*}
Q_\app + \delta Q & = (A_\app + \delta A) \div (B_\app + \delta B) \\
R_\app + \delta R & = (A_\app + \delta A) \mod (B_\app + \delta B)
\end{align*}
where $Q_\app = A_\app \div B_\app$ and $R_\app = A_\app \mod B_\app$.
We have to show that $\NF(\delta Q) \geq \varphi \nfdiv \NF(B)$ and
$\NF(\delta R) \geq \varphi \nfmod \NF(B)$.
Set $\delta X = \delta A - Q_\app \delta B$. 
Using Theorem~\ref{theo:EDivisionNP}, we obtain $\NF(Q_\app) \geq
\NF(A) \nfdiv \NF(B)$ and consequently $\NF(\delta X) \geq \varphi$.
On the other hand, an easy computation yields
$$\delta X = (B + \delta B) \cdot \delta Q + \delta R$$
so that $\delta Q = \delta X \div (B + \delta B)$ and 
$\delta R = \delta X \mod (B + \delta B)$. Using again 
Theorem~\ref{theo:EDivisionNP}, we get the desired result.
\end{proof}

@@

\noindent
{\bf Lattice precision.}
We now move to lattice precision. We pick $A$ and $B$ two polynomials 
of respective degree $n$ and $d$ and assume that they are known at 
precision $O(H_A)$ and $O(H_B)$ respectively:
$$A = A_\app + O(H_A)
\quad \text{and} \quad
B = B_\app + O(H_B).$$
where $H_A \in K_{\leq n}[X]$ and $H_B \in K_{\leq d}[X]$ are lattices. 
According to the results of \cite{padicprec}, in order to determine the 
precision on $A \div B$ and $A \mod B$, we need to compute the 
differential of the mappings $(X,Y) \mapsto X \div Y$ and $(X,Y)
\mapsto X \mod Y$ at the point $(A_\app, B_\app)$. Writing $Q_\app = A_\app \div
B_\app$ and $R_\app = A_\app \mod B_\app$, this can be done by
expanding the relation:
$$A_\app + dA = (B_\app + dB) (Q_\app + dQ) + (R_\app + dR)$$
and neglecting the terms of order $\geq 2$. We get this way 
$dA = B_\app dQ + Q_\app dB + dR$
meaning that $dQ$ and $dR$ appears respectively as the quotient and
the remainder of the Euclidean division of $dX = dA - B_\app dQ$ by $B_\app$.

@@

For this example, we work over $W = \Z_2$ and fix a monic polynomial $M 
\in \Z[X]$ (known exactly) of degree $5$. Our aim is to compare the 
numerical stability of the multiplication in the quotient $\Z_2[X]/M$ 
depending on the precision model we are using. In order to do so, we 
pick $n$ random polynomials $P_1, \ldots, P_n$ in $\Z_2[X]/M(X)$ 
(according to the Haar measure) whose coefficients are all known at 
precision $O(2^N)$ for some large integer $N$. We then compute the 
product of the $P_i$'s using the following quite naive algorithm.

@@

\noindent $\bullet$
in the case of jagged and Newton precision, the precision on the output 
may be written into the form $\sum_{i=0}^4 O(2^{N_i}) \: X^i$ and 
$G = \sum_{i=0}^4 (N_i - N)$;

@@

We define the sequences $(A_i)$ and $(V_i)$ by the recurrence: 
\begin{align*} 
A_0 & = A_\app, \quad V_{0} = V_\app \medskip \\ A_{i+1} 
&= A_i + \big(V_i (P_\app + \delta_P) \:\mod\: A_i\big), \smallskip \\ 
V_{i+1} &= (2 V_i -V_i^2 B_{i+1} ) \mod A_{i+1} \\ 
&  \hspace{1cm}\text{where } B_{i+1} = (P_\app + \delta_P) \div A_{i+1}. 
\end{align*} 
where $A_\app$ and $V_\app$ are those related to $P_\app$. Note that 
$A_\app = a_d \cdot A^{(1)}_\app$ if $a_d$ denotes the coefficient of 
$X^d$ in $P_\app$. By Remark~\ref{rem:slope-factor}, we know that the 
sequence $(A_i)$ converges to $A = a_d \cdot A^{(1)}$ and furthermore:
$$\NP(A - A_\app) \geq \varphi_{|[0,d{-}1]} + 
b_\varphi\big((V_\app \cdot \delta P) \mod A_\app\big)$$ 
since $(V_\app P_\app) \mod A_\app = (1 - V_\app B_\app) \mod A_\app = 
0$. Using repeatedly Lemma~\ref{lem:bphib0}, we obtain:
$$b_\varphi\big((V_\app \cdot \delta P) \mod A_\app\big)
  \geq b_0(V_\app) + b_\varphi(\delta P) \\
  \geq b_\varphi(\delta P) \geq \delta.$$
Thus $\NP(A - A_\app) \geq \varphi_{[0,d{-}1]} + \delta$. Dividing
by $a_d$, we find
$\NP(A^{(1)} - A^{(1)}_\app) \geq \varphi_{A^{(1)}}$ and we are done.
\end{proof}

\todo{Answer here the second question.}

@@

\noindent
{\bf Lattice precision.}
The precision datum is given here by a lattice $H_P$ in $K_{\leq n{-}1}
[X]$; we shall then write:
$$P = P_\app + O(H_P)$$
where $P_\app$ is a \emph{monic} approximation of the inexact polynomial 
$P$ we want to factor. We assume from now that $H_P$ is sufficiently 
small so that the Newton polygon of $P$ is well defined. We then can
define the function:
$$F : P_\app + H_P \to K_{=d}[X] \times K_{=n{-}d}[X]$$
mapping a polynomial $P$ to the couple $(A^{(1)}, B^{(1)})$ obtained
from it. We set $(A_\app^{(1)}, B_\app^{(1)}) = F(P_\app)$.

\noindent
We make the following hypothesis \textbf{(H)}:

\noindent
\hfill
\begin{minipage}{8cm}
The lattice $H_P$ is a first order lattice at every point
of $P_\app + H_P$ in the sense of \cite[Definition~3.3]{padicprec},
\emph{i.e.} for all $P \in P_\app + H_P$:

\vspace{-5mm}

$$F(P + H_P) = F(P) + F'(P)(H_P).$$
\end{minipage}
\hfill \null

\bigskip

\noindent
Obviously \textbf{(H)} gives an answer to the first question we have 
raised above: the precision on the couple $A^{(1)}$ is the lattice 
$H_{A^{(1)}}$ defined as projection on the first component of 
$F'(P_\app)(H_P)$. It turns out that it can be computed explicitely as 
shown by the next proposition.

\begin{prop}
\label{prop:precA1}

The application $P \mapsto A^{(1)}$ is of class $C^1$ on $P_\app + H_P$
and its differential at at some point $P$ is the linear mapping:
$$dP \mapsto dA^{(1)} = (V^{(1)} \: dP) \mod A^{(1)}$$
where $(A^{(1)}, B^{(1)}) = F(P)$ and $V^{(1)}$ is the inverse of 
$B^{(1)}$ modulo $A^{(1)}$.
\end{prop}

\begin{proof}
The function $F$ is injective and a left inverse of it is
$G : (A^{(1)},B^{(1)}) \mapsto A^{(1)}B^{(1)}$.
Clearly $G$ is of class $C^1$ and its differential is given by
\begin{equation}
\label{eq:dPdAdB}
(dA^{(1)}, dB^{(1)}) \mapsto dP = A^{(1)} \: dB^{(1)} + B^{(1)} \: dA^{(1)}.
\end{equation}
Thanks to Bézout Theorem, it is invertible as soon as $A^{(1)}$ and 
$B^{(1)}$ are coprime, which is true because $\NP(A^{(1)})$ and 
$\NP(B^{(1)})$ do not shape a common slope. As a consequence $F$ is of 
class $C^1$ and its differential is obtained by inverting 
Eq.~\eqref{eq:dPdAdB}. Reducing modulo $A^{(1)}$, we get $dP \equiv 
B^{(1)} \: dA^{(1)} \pmod {A^{(1)}}$. The claimed result follows after 
having noticed that $dA^{(1)}$ has degree at most $d{-}1$.
\end{proof}

\todo{Answer here the second question. (The Proposition below should
be useful at some point.)}

\begin{prop}
We assume \textbf{(H)}.

\vspace{-2mm}

\begin{enumerate}[(i)]
\item Let $\tilde A^{(1)} \in K_{=d}[X]$ such that
$\tilde A^{(1)} \tilde B^{(1)} \in P_\app + O(H_P)$ with
$\tilde B^{(1)} = P_\app \div \tilde A^{(1)}$.
Then $\tilde A^{(1)} \in A_\app^{(1)} + H_{A^{(1)}}$.
\item Let in addition $\tilde V^{(1)} \in K_{\leq d{-}1}[X]$ such that:
$$\big(\tilde B^{(1)} \tilde V^{(1)} \cdot H_P\big) \mod \tilde A^{(1)}
= H_P \mod \tilde A^{(1)}.$$
Then $H_{A^{(1)}} = \big(\tilde V^{(1)} \cdot H_P\big) \mod \tilde A^{(1)}$.
\end{enumerate}
\end{prop}

\begin{proof}
\todo{Write it.}
\end{proof}
